Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2015-01-15 draft.tex
Update draft and dual-basis.nb

Compared to the usual maximum-likelihood approach, we find for the method of moments:
\begin{enumerate}
  \item The angular observable ${S_i}$ can be determined
  independently of any other observable ${S_j}$. It is therefore
  much more robust to physics assumptions needed to define the full
  likelihood. In particular, this means one does not have to be
  specific regarding the form of new-physics contributions; in fact,
  one does not even need to be able to explicitly formulate the
  likelihood at all.
\item It is superior for a small number of samples $N$. Likelihood
  fits tend to be numerically unstable if lots of parameters need to
  be estimated from sparse data. This is more severe if the mode of
  the likelihood is near the boundary of the physically allowed
  region.\fred{Is there some LHCb note we could cite?} In many of the
  decays of interest, there are only say dozens of events recorded.
\item The estimate is unbiased for any $N$. In contrast, the
  maximum-likelihood estimate has a bias of order
  $1/N$~\cite{Cox:1968}. In practice, one should keep in mind the
  bias-variance trade-off: it is a well known phenomenon that removing
  the bias usually leads to an increase in variance of the sampling
  distribution of the estimator~\cite[sec. 7.3]{James:2006zz}. From a
  Bayesian decision-theory point of view, both contribute similarly to
  the expected loss associated with deciding on just one value of the
  unknown parameter. It is therefore highly questionable to reduce
  only the bias~\cite[sections 13.8,17.2]{jaynes:2003}. In fact, for
  the results discussed below in \refsec{numerics}, the likelihood
  fits --- if they converge --- produce uncertainties (10--15)\%
  smaller than those from the method of moments.
  \item The approximate multivariate Gaussian distribution of $\vecest{S}$
  allows easier and more correct transfer of the information in the
  data to interested theorists for more accurate fits of
  standard-model and new-physics parameters \citeneeded\fred{Our fits
    and others'?} or for more precise predictions of optimized
  observables; e.g.,
  \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,
    Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf} for definitions
  of such optimized observables in $B\to K^*\ell^+\ell^-$ decays,
  \cite{Faller:2013dwa} for application to the decay $B\to
  \pi\pi\ell^-\bar\nu_\ell$, and \cite{Boeer:2014xx} for observables
  in $\Lambda_b\to\Lambda(\to N\pi)\ell^+\ell^-$). While the
  likelihood also approaches a multivariate Gaussian as $K \to
  \infty$, the two methods differ in their utility as input for
  theorists if $\widehat{\vec{S}}$ is not well
  inside the physical region. For example, suppose there are two
  angular observables that are constrained to a triangular region by
  phase-space or unitarity arguments as \fred{Could add a sketch. It
    is supposed to resemble $A_{\text{FB}},F_L$ for $B \to K^* \ell
    \ell$.}
  \begin{equation}
    \label{eq:constr-ex}
    |S_1| + S_2 \le 1, \, S_1 \in [-1,1],\, S_2 \in [0,1] \,.
  \end{equation}
  It may (and often does) happen in practice that $\widehat{\vec{S}}$
  is close or even outside the allowed region such that a significant
  part of the probability mass covers unphysical values. In a Bayesian
  fit, one would take $\mathcal{N}(\vecest{S} | \vec{S},
  \widehat{\Sigma})$ as the sampling distribution of the ``data'', and
  simply set a uniform prior on the triangle in the $\vec{S}$ plane
  defined by \refeq{constr-ex} to have a well defined problem. This
  could be trivially combined with other independent information in a
  global fit. Someone with a different physics model might have to
  consider a
  different physical region, and could incorporate it just as easily.\\

@@

For the selected class of decays, we can describe that problem as follows: The PDF $P$ has a fixed dependence on the dilepton
helicity angle $\theta_1$ and the azimuthal angle $\theta_3$. (See also \refapp{btokpill} for details
on the angular distribution.) However, at the level of decay amplitudes the dimeson system can have an arbitrarily
large total angular momentum $j$; only its third component is restricted to $m = -1,0,+1$. \fred{check}
We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{j}^{(|m|)}$
with respect to the remaining angle $\theta_2$. Using the observables $S_k(\vec{\nu})$, $k=1,\dots,9$, with $\vec{\nu}=(q^2, k^2)$
and as defined in \refapp{partial-waves} \fred{A lot of mixing between appendices C and E}, the expansion reads
\begin{equation}
    S_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{j,|m|}} S_{k,j}(\vec{\nu}) p_{j}^{(|m|)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
\fred{since when is $S_k a fct. of \cos \theta_2$?}
where the the normalization factor $n_{j,|m|}$ is defined in \refeq{legendre-scalar-product}. Within our
example we have (cf. also \refapp{btokpill})
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,\\
        1\,, & k = 3,4,5,7,8,9\,,
    \end{cases}\,.
\end{equation}
The angular observables $S_{k,j}$ defined as coefficients of this expansion in Legendre polynomials
have the merit of a well defined total angular momentum, and thus are physically discriminable.
As a consequence of the orthogonality of the Legendre polynomials, any mismodelling (or rather, lack of modelling of higher
partial-wave observables) \emph{does not affect} the method of moments as discussed in the previous section. To wit,
ignoring further (orthogonal) terms in the PDF cannot affect the mean vector $(\dots, S_k, \dots)$ nor its variance. \fred{yes!}\\

@@

\todo{size of $M$ in relation to the maximum angular momentum of $\eps$}

@@

First, we find no bias in the determinations of the angular observables even for a small number of events.
However, this comes at the expense of a slight increase of the order of $10\%$ for the statistical uncertainty
on the mean.
Second, in the absence of acceptance effects, a determination using the method of moments is correct and unbiased
even when not all aspects of the physical distribution are modelled. This is explicitl shown for the case of
higher partial waveds in multibody final states.
Third, we develop a systematic method for the dertermination of detector acceptance effects that lead to
dilution and mixing of the angular observables. We present a way to calculate the necessary unfolding matrix,
which is only computationally feasible when using the method of moments.\\
Fourth, the fact that the joint distribution of the results is Gaussian allows for an easy transfer of correlation
information to theory analyses.
Last but not least, the resulting distribution arises without additional constraints necessary for the convergence
of some maximum-likelihood fits. Therefore the results can be directly combined with similarly-obtained results from
a different experiment, without the need to remove the constraints of measurements that stem from different experiments.\\

@@

\section{Application to $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ (S-wave and P-wave)}
\label{app:btokstarll}

\todo{rewrite to use Legendre polynomials}

The PDF for the decay $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ --- up to and including P-wave contributions --- has been calculated
for the most general basis of dimension-six $b\to s$ operators. It reads, expressed in terms of the angular observables $\lbrace J_i\rbrace$ \cite{Blake:2012mb,Bobeth:2012vn}
\begin{equation}
    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3) = \frac{1}{\rmdx{\Gamma}/\rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}} = \frac{3}{8\pi} \sum_k \frac{J_{k}(q^2)}{\rmdx{\Gamma} /\, \rmdx{q^2}} f_k(\cos\theta_1, \cos\theta_2, \theta_3)\,,
\end{equation}
where $\theta_1 \equiv \theta_\ell$ is the dilepton helicity angle, and $\theta_2 \equiv \theta_{K}$ is the $\bar{K}\pi$ helicity angle, and $\theta_3 \equiv \phi$ is the azimuthal angle.
The $q^2$-differential decay width reads
\begin{equation}
    \frac{\rmdx{\Gamma}}{\rmdx{q^2}} = \frac{\big(3 J_{1c} - J_{2c}\big) + 2\big(3J_{1s} - J_{2s}\big)}{3}\,.
\end{equation}
The angular functions\footnote{%
    We consistently use the label "$i$" to denote angular function and observables that arise
    from interference between S-wave and P-wave amplitudes. Our convention is related to the convention
    of \cite{Bobeth:2012vn} through $f_{1i} \leftrightarrow f_{1sc}$, $f_{2i} \leftrightarrow f_{2sc}$.
} read \cite{Blake:2012mb,Bobeth:2012vn}
\begin{equation}
\begin{aligned}
    f_{1s} & = \sin^2\theta_2 &
    f_{1c} & = \cos^2\theta_2 &
    f_{1i} & = \cos\theta_2\\
    f_{2s} & = \sin^2\theta_2 \cos 2\theta_1 &
    f_{2c} & = \cos^2\theta_2 \cos 2\theta_1 &
    f_{2i} & = \cos\theta_2 \cos 2\theta_1\\
    f_{3}  & = \sin^2\theta_2\sin^2\theta_1 \cos 2\theta_3 &
    f_{4}  & = \sin 2\theta_2 \sin 2\theta_1 \cos\theta_3 &
    f_{4i} & = \sin\theta_2 \sin 2\theta_1 \cos\theta_3\\
    f_{5}  & = \sin 2\theta_2 \sin \theta_1 \cos\theta_3 & & &
    f_{5i} & = \sin\theta_2 \sin \theta_1 \cos\theta_3\\
    f_{6s} & = \sin^2\theta_2 \cos\theta_1 &
    f_{6c} & = \cos^2\theta_2 \cos\theta_1\\
    f_{7}  & = \sin 2\theta_2 \sin \theta_1 \sin\theta_3 & & &
    f_{7i} & = \sin\theta_2 \sin \theta_1 \sin\theta_3\\
    f_{8}  & = \sin 2\theta_2 \sin 2\theta_1 \sin\theta_3 & & &
    f_{8i} & = \sin\theta_2 \sin 2\theta_1 \sin\theta_3\\
    f_{9}  & = \sin^2\theta_2\sin^2\theta_1 \sin 2\theta_3\\
\end{aligned}
\end{equation}
In terms of the angles $\theta_1$ and $\theta_2$, instead of their cosines, the PDF reads
\begin{equation}
    P(q^2, \theta_1, \theta_2, \theta_3) = \frac{3}{8\pi} \sum_k S_k(q^2) f_k(\cos\theta_1, \cos\theta_2, \theta_3) \sin\theta_1 \sin\theta_2\,,
\end{equation}
where we also define $S_k(q^2) \equiv J_k(q^2) / (\rmdx{\Gamma} /\, \rmdx{q^2})$.\\

The dual basis reads
\begin{equation}
\begin{aligned}
    \dual{f}_{1s} & = \frac{\big( 63 \sin^2\theta_2 - 42 \cos^2\theta_2\big) + \big(  45 \sin^2\theta_2 - 30 \cos^2\theta_2\big)\cos 2\theta_1}{64} \,,\\
    \dual{f}_{1c} & = \frac{\big(-21 \sin^2\theta_2 + 84 \cos^2\theta_2\big) + \big( -15 \sin^2\theta_2 + 60 \cos^2\theta_2\big)\cos 2\theta_1}{32} \,,\\
    \dual{f}_{2s} & = \frac{\big( 45 \sin^2\theta_2 - 30 \cos^2\theta_2\big) + \big(+135 \sin^2\theta_2 - 90 \cos^2\theta_2\big)\cos 2\theta_1}{64} \,,\\
    \dual{f}_{2c} & = \frac{\big(-15 \sin^2\theta_2 + 60 \cos^2\theta_2\big) + \big( -45 \sin^2\theta_2 +180 \cos^2\theta_2\big)\cos 2\theta_1}{32} \,,
\end{aligned}
\end{equation}
as well as
\begin{equation}
\begin{aligned}
    \dual{f}_{1i} & = \frac{\big(21 + 15 \cos1\theta_2\big)\cos\theta_2}{16}               \,,&
    \dual{f}_{2i} & = \frac{\big(15 + 45 \cos1\theta_2\big)\cos\theta_2}{16}               \,,\\
    \dual{f}_{3}  & = \frac{75 \sin^2\theta_2 \sin^2\theta_1 \cos2\theta_3}{32}                \,,& &\\
    \dual{f}_{4}  & = \frac{75 \sin 2\theta_2 \sin 2\theta_1 \cos \theta_3}{32}                \,,&
    \dual{f}_{4i} & = \frac{15 \sin  \theta_2 \sin 2\theta_1 \cos \theta_3}{8}                 \,,\\
    \dual{f}_{5}  & = \frac{15 \sin 2\theta_2 \sin  \theta_1 \cos \theta_3}{8}                 \,,&
    \dual{f}_{5i} & = \frac{3  \sin  \theta_2 \sin  \theta_1 \cos \theta_3}{2}                 \,,\\
    \dual{f}_{6s} & = \frac{\big( 9\sin^2\theta_2 -  6 \cos^2\theta_2\big)\cos\theta_1}{4} \,,&
    \dual{f}_{6c} & = \frac{\big(-3\sin^2\theta_2 + 12 \cos^2\theta_2\big)\cos\theta_1}{2} \,,\\
    \dual{f}_{7}  & = \frac{15 \sin 2\theta_2 \sin  \theta_1 \sin \theta_3}{8}                 \,,&
    \dual{f}_{7i} & = \frac{3  \sin  \theta_2 \sin  \theta_1 \sin \theta_3}{2}                 \,,\\
    \dual{f}_{8}  & = \frac{75 \sin 2\theta_2 \sin 2\theta_1 \sin \theta_3}{32}                \,,&
    \dual{f}_{8i} & = \frac{15 \sin  \theta_2 \sin 2\theta_1 \sin \theta_3}{8}                 \,,\\
    \dual{f}_{9}  & = \frac{75 \sin^2\theta_2 \sin^2\theta_1 \sin2\theta_3}{32}                \,.\\
\end{aligned}
\end{equation}
such that
\begin{equation}
    \int_0^\pi \rmdx{\theta_1} \int_0^\pi \rmdx{\theta_2} \int_0^{2\pi} \rmdx{\theta_3} \dual{f}_k(\theta_1,\theta_2,\theta_3) P(q^2, \theta_1,\theta_2,\theta_3) = S_k(q^2)\,.
\end{equation}

\section{Application to $\bar{B}\to P_1 P_2\ell^+\ell^-$ (all partial waves)}
\label{app:btokpill}

\todo{rewrite to use Legendre polynomials}

The class of semileptonic $\bar{B}\to P_1 P_2 \ell_1\bar\ell_2$ decays, where $P_1$, and $P_2$ are pseudoscalar mesons,
and $\ell_1$ and $\bar\ell_2$ are charged or neutral leptons, includes decays
such as $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ and $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$.
Their decay PDFs can generally be written as \cite{Lee:1992ih}
\begin{equation}
    P(q^2, p^2, \cos\theta_1, \cos \theta_2, \theta_3) = \frac{1}{\langle \Gamma\rangle}\frac{\rmdx{^5 \Gamma}}{\rmdx{q^2} \rmdx{p^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}} = \frac{1}{4\pi} \sum_k \frac{I_k(q^2, p^2, \cos\theta_2)}{\Gamma} f_k(\cos\theta_1,\theta_3)\,,
\end{equation}
where
\begin{equation}
\begin{aligned}
    f_1 & = 1\,,              &
    f_2 & = \cos 2\theta_1\,, &
    f_3 & = \sin^2\theta_1 \cos 2\theta_3\,,\\
    f_4 & = \sin 2\theta_1 \cos  \theta_3\,,&
    f_5 & = \sin  \theta_1 \cos  \theta_3\,,&
    f_6 & = \cos  \theta_1\,, \\
    f_7 & = \sin  \theta_1 \sin  \theta_3\,,&
    f_8 & = \sin 2\theta_1 \sin  \theta_3\,,&
    f_9 & = \sin^2\theta_1 \sin 2\theta_3\,.\\
\end{aligned}
\end{equation}
Here we denote dilepton mass squared and the dilepton helicity angle as $q^2$ and $\theta_1$, while
the dimeson mass squared and the dimeson helicity angles are denoted as $p^2$ and $\theta_2$. The
azimuthal angle is denoted as $\theta_3$.
The total decay width is
\begin{equation}
    \langle \Gamma\rangle = \int_0^\pi \rmdx{\theta_2} \rmdx{q^2} \rmdx{p^2} \sin\theta_2 \frac{3I_1(q^2, p^2, \cos\theta_2) - I_2(q^2, p^2, \cos\theta_2)}{3}\,.
\end{equation}
As previous, we transform from a $\cos\theta_2$-dependence to a $\theta_2$-dependence.
\begin{equation}
\begin{aligned}
    P(q^2, p^2, \theta_1, \theta_2, \theta_3)
    & = \frac{1}{4\pi} \sum_k \frac{I_k(q^2, p^2, \cos\theta_2)}{\Gamma} f_i(\theta_1, \theta_3) \sin\theta_1 \sin\theta_2\\
    & \equiv \frac{1}{4\pi} \sum_k S_k(q^2, p^2, \theta_2) f_k(\theta_1, \theta_3) \sin\theta_1\,.
\end{aligned}
\end{equation}
where we also define $S_k(q^2, p^2, \theta_2) \equiv \sin\theta_2 I_k(q^2, p^2, \cos\theta_2) / \langle \Gamma\rangle$.\\

The dual basis reads
\begin{equation}
\begin{aligned}
    \dual{f}_1 & = \frac{21 + 15 \cos 2\theta_1}{16}\,, &
    \dual{f}_2 & = \frac{15 + 45 \cos 2\theta_1}{16}\,, &
    \dual{f}_3 & = \frac{15}{4} \sin^2\theta_1 \cos 2\theta_3     \,,\\
    \dual{f}_4 & = \frac{15}{4} \sin 2\theta_1 \cos  \theta_3     \,, &
    \dual{f}_5 & = 3            \sin  \theta_1 \cos  \theta_3     \,, &
    \dual{f}_6 & = 3 \cos\theta_1                   \,,\\
    \dual{f}_7 & = 3            \sin  \theta_1 \sin  \theta_3     \,, &
    \dual{f}_8 & = \frac{15}{4} \sin 2\theta_1 \sin  \theta_3     \,, &
    \dual{f}_9 & = \frac{15}{4} \sin^2\theta_1 \sin 2\theta_3     \,.\\
\end{aligned}
\end{equation}
such that
\begin{equation}
    \int_0^\pi \rmdx{\theta_1} \int_0^{2\pi} \rmdx{\theta_3} \dual{f}_i(\theta_1,\theta_3) P(q^2, k^2, \theta_1,\theta_2,\theta_3) = S_i(q^2, k^2, \theta_2)\,.
\end{equation}

@@

The PDF for the decay --- in the presence of Standard-Model operators and their chirality-flipped counter parts --- reads \cite{Boeer:2014xx}
\begin{equation}
    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3) = \frac{1}{\rmdx{\Gamma} /\, \rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}} = \sum_i S_i f_i(\cos\theta_1, \cos\theta_2, \theta_3)\,,
\end{equation}
where $q^2$ denotes the dilepton mass squared, $\theta_1 \equiv \theta_\ell$ and $\theta_2 \equiv \theta_\Lambda$ denote the
helicity angles in the dilepton and
$N\pi$ systems, respectively, and $\theta_3 = \phi$ denotes the azimuthal angle.
The index $i$ should be interpreted as a multi-index, $i \equiv (l_1, l_2, m)$,
where $0 \leq l_1 \leq 2$ and $0 \leq l_2 \leq 1$ denote the total angular momentum in the
dilepton and the $N\pi$ system, respectively, and $-1 \leq m \leq 1$
is the third component of either of the angular momenta.
Our choice of an orthonormal basis and its dual reads
\begin{equation}
\begin{aligned}
    f_{l_1, l_2, m}(\cos\theta_1, \cos\theta_2, \theta_3)
    & = \frac{1}{8\pi} \sqrt{\frac{(l_1 - |m|)!\,(l_2 - |m|)!}{(l_1 + |m|)!\,(l_2 + |m|)!}} p_{l_1}^{|m|}(\cos\theta_1) p_{l_2}^{|m|}(\cos\theta_2)\\
    & \times \begin{cases}
            \cos(|m|\theta_3) & m > 0\\
            \sin(|m|\theta_3) & m < 0\\
            1                 & m = 0
        \end{cases}\,,\\
    \tilde{f}_{l_1, l_2, m}(\cos\theta_1, \cos\theta_2, \theta_3)
    & = (2l_1 + 1)(2l_2 + 1)\sqrt{\frac{(l_1 - m)!\,(l_2 - m)!}{(l_1 + m)!\,(l_2 + m)!}} p_{l_1}^m(\cos\theta_1) p_{l_2}^m(\cos\theta_2)\\
    & \times \begin{cases}
            2 \cos(|m|\theta_3) & m > 0\\
            2 \sin(|m|\theta_3) & m < 0\\
            1                   & m = 0
        \end{cases}\,.\\
\end{aligned}
\end{equation}
The correspondence between out choice of angular observables in the angular momentum basis, and the angular
observables as defined in reference \cite{Boeer:2014xx} reads
\begin{equation}
\begin{aligned}
    S_{0, 0,  0} & = 1\,,                                                             &
    S_{1, 0,  0} & = \frac{3 K_{1c}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       &
    S_{2, 0,  0} & = \frac{2(K_{1cc} - K_{1ss})}{\rmdx\Gamma/\rmdx{q^2}}\,,           \\
    S_{0, 1,  0} & = \frac{K_{2cc} + 2 K_{2ss}}{\rmdx\Gamma/\rmdx{q^2}}\,,            &
    S_{1, 1,  0} & = \frac{3 K_{2c}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       &
    S_{2, 1,  0} & = \frac{2(K_{2cc} - K_{2ss})}{\rmdx\Gamma/\rmdx{q^2}}\,,           \\
                 &                                                                    &
    S_{1, 1, -1} & = \frac{6 K_{4s}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       &
    S_{2, 1, -1} & = \frac{2 \sqrt{3} K_{4sc}}{\rmdx\Gamma/\rmdx{q^2}}\,,             \\
                 &                                                                    &
    S_{1, 1, +1} & = \frac{6 K_{3s}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       &
    S_{2, 1, +1} & = \frac{2 \sqrt{3} K_{3sc}}{\rmdx\Gamma/\rmdx{q^2}}\,,
\end{aligned}
\end{equation}
where the decay width is
\begin{equation}
    \frac{\rmdx{\Gamma}} {\rmdx{q^2}} = 2 K_{1ss} + K_{1cc}\,.
\end{equation}

@@

>>>>

Compared to the usual maximum-likelihood approach, we find for the method of moments:
\begin{enumerate}
  \item The angular observable ${S_i}$ can be determined
  independently of any other observable ${S_j}$. It is therefore
  much more robust to physics assumptions needed to define the full
  likelihood. In particular, this means one does not have to be
  specific regarding the form of new-physics contributions; in fact,
  one does not even need to be able to explicitly formulate the
  likelihood at all.
\item It is superior for a small number of samples $N$. Likelihood
  fits tend to be numerically unstable if lots of parameters need to
  be estimated from sparse data. This is more severe if the mode of
  the likelihood is near the boundary of the physically allowed
  region.\fred{Is there some LHCb note we could cite?} In many of the
  decays of interest, there are only say dozens\todo{check!} of events recorded.
\item The estimate is unbiased for any $N$. In contrast, the
  maximum-likelihood estimate has a bias of order
  $1/N$~\cite{Cox:1968}. In practice, one should keep in mind the
  bias-variance trade-off: it is a well known phenomenon that removing
  the bias usually leads to an increase in variance of the sampling
  distribution of the estimator~\cite[sec. 7.3]{James:2006zz}. From a
  Bayesian decision-theory point of view, both contribute similarly to
  the expected loss associated with deciding on just one value of the
  unknown parameter. It is therefore highly questionable to reduce
  only the bias~\cite[sections 13.8,17.2]{jaynes:2003}. In fact, for
  the results discussed below in \refsec{numerics}, the likelihood
  fits --- if they converge --- produce uncertainties (10--15)\%
  smaller than those from the method of moments.
  \item The approximate multivariate Gaussian distribution of $\vecest{S}$
  allows easier and more correct transfer of the information in the
  data to interested theorists for more accurate fits of
  standard-model and new-physics parameters \citeneeded\fred{Our fits
    and others'?} or for more precise predictions of optimized
  observables; e.g.,
  \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,
    Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf} for definitions
  of such optimized observables in $B\to K^*\ell^+\ell^-$ decays,
  \cite{Faller:2013dwa} for application to the decay $B\to
  \pi\pi\ell^-\bar\nu_\ell$, and \cite{Boeer:2014xx} for observables
  in $\Lambda_b\to\Lambda(\to N\pi)\ell^+\ell^-$). While the
  likelihood also approaches a multivariate Gaussian as $K \to
  \infty$, the two methods differ in their utility as input for
  theorists if $\widehat{\vec{S}}$ is not well
  inside the physical region. For example, suppose there are two
  angular observables that are constrained to a triangular region by
  phase-space or unitarity arguments as \fred{Could add a sketch. It
    is supposed to resemble $A_{\text{FB}},F_L$ for $B \to K^* \ell
    \ell$.}
  \begin{equation}
    \label{eq:constr-ex}
    |S_1| + S_2 \le 1, \, S_1 \in [-1,1],\, S_2 \in [0,1] \,.
  \end{equation}
  It may (and often does) happen in practice that $\widehat{\vec{S}}$
  is close or even outside the allowed region such that a significant
  part of the probability mass covers unphysical values. In a Bayesian
  fit, one would take $\mathcal{N}(\vecest{S} | \vec{S},
  \widehat{\Sigma})$ as the sampling distribution of the ``data'', and
  simply set a uniform prior on the triangle in the $\vec{S}$ plane
  defined by \refeq{constr-ex} to have a well defined problem. This
  could be trivially combined with other independent information in a
  global fit. Someone with a different physics model might have to
  consider a
  different physical region, and could incorporate it just as easily.\\

@@

For the selected class of decays, we can describe that problem as follows: The PDF $P$ has a fixed dependence on the dilepton
helicity angle $\theta_1$ and the azimuthal angle $\theta_3$. (See also \refapp{btokpill} for details
on the angular distribution.) However, at the level of decay amplitudes the dimeson system can have an arbitrarily
large total angular momentum $j$; only its third component is restricted to $m = -1,0,+1$.
It is then usual to compute the angular distribution explicitly in terms of $\theta_1$ and $\theta_3$, but leave
the angular observables dependent on the remaining helicity angle $\theta_2$.
We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{j}^{(|m|)}$
with respect to the remaining angle $\theta_2$. Using the observables $S_k(\vec{\nu})$, $k=1,\dots,9$, with $\vec{\nu}=(q^2, k^2)$
and as defined in \refapp{partial-waves} \fred{A lot of mixing between appendices C and E}, the expansion reads
\begin{equation}
    S_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{j,|m|}} S_{k,j}(\vec{\nu}) p_{j}^{(|m|)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where the the normalization factor $n_{j,|m|}$ is defined in \refeq{legendre-scalar-product}. Within our
example we have (cf. also \refapp{btokpill})
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,\\
        1\,, & k = 3,4,5,7,8,9\,,
    \end{cases}\,.
\end{equation}
The angular observables $S_{k,j}$ defined as coefficients of this expansion in Legendre polynomials
have the merit of a well defined total angular momentum, and thus are physically discriminable.
As a consequence of the orthogonality of the Legendre polynomials, any mismodelling (or rather, lack of modelling of higher
partial-wave observables) \emph{does not affect} the method of moments as discussed in the previous section. To wit,
ignoring further (orthogonal) terms in the PDF cannot affect the mean vector $(\dots, S_k, \dots)$ nor its variance. \fred{yes!}\\

@@

It remains to be emphasized that the dimensionality of the raw angular observables is not completely
fixed by the number of angular momentum configurations of the process at hand. Let us revisit the
decay $B\to K\ell^+\ell^-$, with only one angle,  to clear this up. For $B\to K\ell^+\ell^-$, the
total angular momentum associated the dilepton pair can, at the level of observables, be either $0$,
$1$ or $2$. We assume now that the acceptance function $\eps$ can be conveniently expanded in terms
of Legendre polynomials up to maximal angular momentum $L$. From angular momentum conservation
follows immediately that $\dim Q = L + 2$, and thus the unfolding matrix $M$ will be of
dimension $L + 2 \times L + 2$. (For a multiangular distribution the maximum angular momentum
has to be determined in the same way for each angle.)
Since the $\eps$ depends highly on the nature of the particle detector at hand, it remains for
the experimental collaborations to determine the maximum order of such an expansion.

@@

First, we find no bias in the determinations of the angular observables even for a small number of events.
However, this comes at the expense of a slight increase of the order of $10\%$\todo{check!} for the statistical uncertainty
on the mean.
Second, in the absence of acceptance effects, a determination using the method of moments is correct and unbiased
even when not all aspects of the physical distribution are modelled. This is explicitl shown for the case of
higher partial waveds in multibody final states.
Third, we develop a systematic method for the dertermination of detector acceptance effects that lead to
dilution and mixing of the angular observables. We present a way to calculate the necessary unfolding matrix,
which is only computationally feasible when using the method of moments.\\
Fourth, the fact that the joint distribution of the results is Gaussian allows for an easy transfer of correlation
information to theory analyses.
Last but not least, the resulting distribution arises without additional constraints necessary for the convergence
of some maximum-likelihood fits. Therefore the results can be directly combined with similarly-obtained results from
a different experiment, without the need to remove the constraints of measurements that stem from different experiments.\\

@@

@@

The PDF for the decay --- in the presence of Standard-Model operators and their chirality-flipped counter parts --- reads \cite{Boeer:2014xx}
\begin{equation}
    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3) = \frac{1}{\rmdx{\Gamma} /\, \rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}} = \sum_i S_i f_i(\cos\theta_1, \cos\theta_2, \theta_3)\,,
\end{equation}
where $q^2$ denotes the dilepton mass squared, $\theta_1 \equiv \theta_\ell$ and $\theta_2 \equiv \theta_\Lambda$ denote the
helicity angles in the dilepton and
$N\pi$ systems, respectively, and $\theta_3 = \phi$ denotes the azimuthal angle.
The index $i$ should be interpreted as a multi-index, $i \equiv (l_1, l_2, m)$,
where $0 \leq l_1 \leq 2$ and $0 \leq l_2 \leq 1$ denote the total angular momentum in the
dilepton and the $N\pi$ system, respectively, and $-1 \leq m \leq 1$
is the third component of either of the angular momenta.
Our choice of an orthonormal basis and its dual reads
\begin{equation}
\label{eq:lambdab:bases}
\begin{aligned}
    f_{l_1, l_2, m}(\cos\theta_1, \cos\theta_2, \theta_3)
    & = \sqrt{\frac{(l_1 - |m|)!\,(l_2 - |m|)!}{(l_1 + |m|)!\,(l_2 + |m|)!}} p_{l_1}^{|m|}(\cos\theta_1) p_{l_2}^{|m|}(\cos\theta_2)\\
    & \times \begin{cases}
            \cos(|m|\theta_3) & m > 0\\
            \sin(|m|\theta_3) & m < 0\\
            1                 & m = 0
        \end{cases}\,,\\
    \tilde{f}_{l_1, l_2, m}(\cos\theta_1, \cos\theta_2, \theta_3)
    & = \frac{(2l_1 + 1)(2l_2 + 1)}{8\pi}\sqrt{\frac{(l_1 - m)!\,(l_2 - m)!}{(l_1 + m)!\,(l_2 + m)!}} p_{l_1}^m(\cos\theta_1) p_{l_2}^m(\cos\theta_2)\\
    & \times \begin{cases}
            2 \cos(|m|\theta_3) & m > 0\\
            2 \sin(|m|\theta_3) & m < 0\\
            1                   & m = 0
        \end{cases}\,.\\
\end{aligned}
\end{equation}
The correspondence between out choice of angular observables in the angular momentum basis, and the angular
observables as defined in reference \cite{Boeer:2014xx} reads
\begin{equation}
\begin{aligned}
    8\pi S_{0, 0,  0} & = 1\,,                                                             &
    8\pi S_{1, 0,  0} & = \frac{3 K_{1c}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       &
    8\pi S_{2, 0,  0} & = \frac{2(K_{1cc} - K_{1ss})}{\rmdx\Gamma/\rmdx{q^2}}\,,           \\
    8\pi S_{0, 1,  0} & = \frac{K_{2cc} + 2 K_{2ss}}{\rmdx\Gamma/\rmdx{q^2}}\,,            &
    8\pi S_{1, 1,  0} & = \frac{3 K_{2c}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       &
    8\pi S_{2, 1,  0} & = \frac{2(K_{2cc} - K_{2ss})}{\rmdx\Gamma/\rmdx{q^2}}\,,           \\
                 &                                                                    &
    8\pi S_{1, 1, -1} & = \frac{6 K_{4s}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       &
    8\pi S_{2, 1, -1} & = \frac{2 \sqrt{3} K_{4sc}}{\rmdx\Gamma/\rmdx{q^2}}\,,             \\
                 &                                                                    &
    8\pi S_{1, 1, +1} & = \frac{6 K_{3s}}{\rmdx\Gamma/\rmdx{q^2}}\,,                       &
    8\pi S_{2, 1, +1} & = \frac{2 \sqrt{3} K_{3sc}}{\rmdx\Gamma/\rmdx{q^2}}\,,
\end{aligned}
\end{equation}
where the decay width is
\begin{equation}
    \frac{\rmdx{\Gamma}} {\rmdx{q^2}} = 2 K_{1ss} + K_{1cc}\,.
\end{equation}

@@

For the purpose of unfolding acceptance effects as laid down in \refsec{systematics:acceptance}, it is instrumental
to know that $f_{0,0,0} \equiv 1$, and that $\max_{\cos\theta_1,\cos\theta_2,\phi} |f_{l_1, l_2, m}| < 1$.
The recipe's generating PDFs are therefore
\begin{equation}
    P(x|\lbrace S_i^{(j)}\rbrace)\,,\quad\text{with}\quad S_i^{(j)} = \frac{1}{8\pi}\left(\delta_{i,(0,0,0)} + \delta_{i,j}\right)\,,
\end{equation}
where $j$ ís now also a multi-index representing $j \equiv (\tilde l_1, \tilde l_2, \tilde m)$.

\section{Application to $\bar{B}\to\bar{K}\pi\ell^+\ell^-$}
\label{app:btokstarll}
